---
title: Benchmark of stringr replacement
output: 
  github_document:
    toc: true
---

# Context

As part of [#1549](https://github.com/yihui/knitr/issues/1549), we removed **stringr** dependency by replacing all functions used with base R equivalent.

This document aim to be a benchmark of each replacement. We'll use [**bench**](https://bench.r-lib.org/) to do the benchmark, and compare **stringr** function with its replacement. Each benchmark will be done in the context of there initial usage in **knitr**.

## Session info

```{r}
xfun::session_info(c("knitr", "bench", "stringr"))
```

## Helpers

```{r}
knitr_example = function(...) system.file('examples', ..., package = 'knitr')
```

# Benchmarking

## `stringr::str_replace_all()`

| Commit                                                                                     | PR                                                | File                                                                                                                            |
|--------------------------------------------------------------------------------------------|---------------------------------------------------|---------------------------------------------------------------------------------------------------------------------------------|
| [9c92eff1](https://github.com/yihui/knitr/commit/9c92eff1f22b1e6c3bf5dc538dd8c0a2edc8ad48) | [#2174](https://github.com/yihui/knitr/pull/2174) | [utils-vignettes.R#L158](https://github.com/yihui/knitr/blob/9c92eff1f22b1e6c3bf5dc538dd8c0a2edc8ad48/R/utils-vignettes.R#L158) |

Before:

``` r
x[!i] = stringr::str_replace_all(x[!i], p$inline.code, '') # remove inline code
```

After:

``` r
x[!i] = gsub(p$inline.code, '', x[!i], perl = TRUE) # remove inline code
```

```{r}
file = knitr_example('knitr-minimal.Rnw')
x = xfun::read_utf8(file)
p = knitr:::detect_pattern(x, tolower(xfun::file_ext(file)))
p = knitr:::all_patterns[[p]]
p1 = p$chunk.begin; p2  = p$chunk.end
m = knitr:::group_indices(grepl(p1, x), grepl(p2, x))
i = m %% 2 == 0
```

```{r}
res = bench::mark(
  stringr = stringr::str_replace_all(x[!i], p$inline.code, ''),
  new = gsub(p$inline.code, '', x[!i], perl = TRUE),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

## `stringr::str_trim()`

+-------------------------------------------------------------------------------------------+---------------------------------------------------+---------------------------------------------------------------------------------------------------------------+
| Commit                                                                                    | PR                                                | File                                                                                                          |
+===========================================================================================+===================================================+===============================================================================================================+
| [8fa7d17](https://github.com/yihui/knitr/commit/8fa7d1710bae72f9636ef64fa9286669329f1a9c) | [#2177](https://github.com/yihui/knitr/pull/2177) | [parser.R#L491](https://github.com/yihui/knitr/blob/8fa7d1710bae72f9636ef64fa9286669329f1a9c/R/parser.R#L491) |
|                                                                                           |                                                   |                                                                                                               |
|                                                                                           |                                                   | [utils.R#L77](https://github.com/yihui/knitr/blob/8fa7d1710bae72f9636ef64fa9286669329f1a9c/R/utils.R#L77)     |
+-------------------------------------------------------------------------------------------+---------------------------------------------------+---------------------------------------------------------------------------------------------------------------+

Before:

``` r
labels = stringr::str_trim(gsub(lab, '\\3', sapply(groups, `[`, 1)))
```

After:

``` r
x[!i] = gsub(p$inline.code, '', x[!i], perl = TRUE) # remove inline code
```

```{r}
path = "https://raw.githubusercontent.com/yihui/knitr-examples/46c8d1db0cf0c9ab04432444079927324c4c3688/113-foo.R"
lines = xfun::read_utf8(path)
lab = knitr:::.sep.label
idx = cumsum(grepl(lab, lines))
groups = unname(split(lines, idx))
```

```{r}
content = gsub(lab, '\\3', sapply(groups, `[`, 1))

res = bench::mark(
  stringr = stringr::str_trim(content),
  new = trimws(content),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

| Commit                                                                                    | PR                                                | File                                                                                                      |
|-------------------------------------------------------------------------------------------|---------------------------------------------------|-----------------------------------------------------------------------------------------------------------|
| [8fa7d17](https://github.com/yihui/knitr/commit/8fa7d1710bae72f9636ef64fa9286669329f1a9c) | [#2177](https://github.com/yihui/knitr/pull/2177) | [utils.R#L77](https://github.com/yihui/knitr/blob/8fa7d1710bae72f9636ef64fa9286669329f1a9c/R/utils.R#L77) |

Before:

``` r
stringr::str_trim(stringr::str_split(string, ';|,')[[1]])
```

After:

``` r
trimws(stringr::str_split(string, ';|,')[[1]])
```

```{r}
string = ' .5,.6 , .7; .9 '
splitted = stringr::str_split(string, ';|,')[[1]]

res = bench::mark(
  stringr = stringr::str_trim(splitted),
  new = trimws(splitted),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

## `stringr::str_dup()`

| Commit                                                                                    | PR                                                | File                                                                                                            |
|-------------------------------------------------------------------------------------------|---------------------------------------------------|-----------------------------------------------------------------------------------------------------------------|
| [1ce8286](https://github.com/yihui/knitr/commit/1ce82862b689338ce275269056ebfcc96eb975ca) | [#2186](https://github.com/yihui/knitr/pull/2186) | [parser.R#L3488](https://github.com/yihui/knitr/blob/1ce82862b689338ce275269056ebfcc96eb975ca/R/parser.R#L3488) |

Before:

``` r
cat(' ', stringr::str_dup('~', getOption('width') - 10L), '\n')
```

After:

``` r
rep_str = function(x, n, sep = '') paste(rep(x, n), collapse = sep)
cat(' ', rep_str('~', getOption('width') - 10L), '\n')
```

```{r}
res = bench::mark(
  stringr = stringr::str_dup('~', getOption('width') - 10L),
  new = paste(rep('~', getOption('width') - 10L), collapse = ''),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

## `stringr::str_split()`

+------------------------------------------------------------------------------------------+---------------------------------------------------+---------------------------------------------------------------------------------------------------------------+
| Commit                                                                                   | PR                                                | File                                                                                                          |
+==========================================================================================+===================================================+===============================================================================================================+
| [67b973](https://github.com/yihui/knitr/commit/67b973dc87c9d479cd5fbd76806e3a91f5e7d18c) | [#2187](https://github.com/yihui/knitr/pull/2187) | [pandoc.R#L125](https://github.com/yihui/knitr/blob/67b973dc87c9d479cd5fbd76806e3a91f5e7d18c/R/pandoc.R#L125) |
|                                                                                          |                                                   |                                                                                                               |
|                                                                                          |                                                   | [utils.R#L77](https://github.com/yihui/knitr/blob/67b973dc87c9d479cd5fbd76806e3a91f5e7d18c/R/utils.R#L77)     |
+------------------------------------------------------------------------------------------+---------------------------------------------------+---------------------------------------------------------------------------------------------------------------+

Before:

``` r
x = stringr::str_split(x, '\n')
```

After:

``` r
# patch strsplit() to split '' into '' instead of character(0)
str_split = function(x, split, ...) {
  y = strsplit(x, split, ...)
  y[x == ''] = list('')
  y
}

x = str_split(x, '\n')
```

```{r}
con = url("https://raw.githubusercontent.com/yihui/knitr-examples/46c8d1db0cf0c9ab04432444079927324c4c3688/084-pandoc.pandoc")
x = read.dcf(con)
close(con)
x = x[!is.na(x)]
```

```{r}
str_split = function(x, split, ...) {
  y = strsplit(x, split, ...)
  y[x == ''] = list('')
  y
}
res = bench::mark(
  stringr = stringr::str_split(x, '\n'),
  new = str_split(x, '\n'),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

## `stringr::str_sub()`

+-------------------------------------------------------------------------------------------+---------------------------------------------------+-------------------------------------------------------------------------------------------------------------------+
| Commit                                                                                    | PR                                                | File                                                                                                              |
+===========================================================================================+===================================================+===================================================================================================================+
| [5a2cb72](https://github.com/yihui/knitr/commit/5a2cb72f63aa64b757418df99fc2ad3ef5292ac0) | [#2195](https://github.com/yihui/knitr/pull/2195) | [utils-string.R](https://github.com/yihui/knitr/blob/5a2cb72f63aa64b757418df99fc2ad3ef5292ac0/R/utils-string.R)\  |
|                                                                                           |                                                   | [block.R#L564](https://github.com/yihui/knitr/blob/5a2cb72f63aa64b757418df99fc2ad3ef5292ac0/R/block.R#L564)\      |
|                                                                                           |                                                   | [utils.R#L92-L93](https://github.com/yihui/knitr/blob/5a2cb72f63aa64b757418df99fc2ad3ef5292ac0/R/utils.R#L92-L93) |
+-------------------------------------------------------------------------------------------+---------------------------------------------------+-------------------------------------------------------------------------------------------------------------------+

Before:

``` r
stringr::str_sub(input, loc[i, 1], loc[i, 2]) = if (length(res)) {
      paste(hook(res), collapse = '')
    } else ''
```

After:

``` r
# replace parts of a string with new values; `pos` is a matrix of positions and
# each row is a pair of [start, end]
str_replace = function(x, pos, value) {
  if (length(x) != 1) stop("Only a character scalar is supported.")
  # extract parts of the string that are outside [start, end]
  m = rbind(pos[, 1] - 1, pos[, 2] + 1)
  m = matrix(c(1, m, nchar(x)), nrow = 2)
  y = substring(x, m[1, ], m[2, ])
  paste(rbind(y, c(value, '')), collapse = '')
}

str_replace(input, block$location, ans)
```

```{r}
input = "inline `r (function() 1)()`"
location = matrix(c(8,27), ncol = 2, byrow = TRUE)
ans = "1"

str_replace = function(x, pos, value) {
  if (length(x) != 1) stop("Only a character scalar is supported.")
  # extract parts of the string that are outside [start, end]
  m = rbind(pos[, 1] - 1, pos[, 2] + 1)
  m = matrix(c(1, m, nchar(x)), nrow = 2)
  y = substring(x, m[1, ], m[2, ])
  paste(rbind(y, c(value, '')), collapse = '')
}

# wrapper to approximate what is done internally
old_replace = function(input, loc, ans) {
  stringr::str_sub(input, loc[1, 1], loc[1, 2]) = ans
  return(input)
}
```

```{r}
res = bench::mark(
  stringr = old_replace(input, location, ans),
  new = str_replace(input, location, ans),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

+-------------------------------------------------------------------------------------------+---------------------------------------------------+----------------------------------------------------------------------------------------------------------------------+
| Commit                                                                                    | PR                                                | File                                                                                                                 |
+===========================================================================================+===================================================+======================================================================================================================+
| [5a2cb72](https://github.com/yihui/knitr/commit/5a2cb72f63aa64b757418df99fc2ad3ef5292ac0) | [#2195](https://github.com/yihui/knitr/pull/2195) | [header.R#L59-L60](https://github.com/yihui/knitr/blob/5a2cb72f63aa64b757418df99fc2ad3ef5292ac0/R/header.R#L59-L60)\ |
|                                                                                           |                                                   | [header.R#L90-L91](https://github.com/yihui/knitr/blob/5a2cb72f63aa64b757418df99fc2ad3ef5292ac0/R/header.R#L90-L91)  |
+-------------------------------------------------------------------------------------------+---------------------------------------------------+----------------------------------------------------------------------------------------------------------------------+

Before:

``` r
tmp = stringr::str_sub(doc[i], l[, 1], l[, 2])
stringr::str_sub(doc[i], l[,1], l[,2]) = paste0(tmp, '\n', make_header_html())
```

After :

``` r
tmp = substr(doc[i], l[, 1], l[, 2])
doc[i] = str_replace(doc[i], l, paste0(tmp, '\n', make_header_html()))
```

```{r}
b = knitr::all_patterns$html$header.begin
doc = c("<head>", "<!--content-->", "</head>", "<body>", "</body>")
i = grep(b, doc)
l = stringr::str_locate(doc[i], b)
```

```{r}
res = bench::mark(
  stringr = stringr::str_sub(doc[i], l[, 1], l[, 2]),
  new = substr(doc[i], l[, 1], l[, 2]),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

```{r}
tmp = stringr::str_sub(doc[i], l[, 1], l[, 2])
header = knitr:::make_header_html()

old_fun = function(doc, l, tmp, header) {
  stringr::str_sub(doc[i], l[,1], l[,2]) = paste0(tmp, header)
  doc
}

new_fun = function(doc, l, tmp, header) {
  doc[i] = str_replace(doc[i], l, paste0(tmp, header))
  doc
}

res = bench::mark(
  stringr = old_fun(doc, l, tmp, header),
  new = new_fun(doc, l, tmp, header),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

Before:

``` r
tmp = stringr::str_sub(doc[i], l[, 1], l[, 2])
stringr::str_sub(doc[i], l[,1], l[,2]) = paste0(tmp, make_header_latex(doc))
```

After :

``` r
tmp = substr(doc[i], l[, 1], l[, 2])
doc[i] = str_replace(doc[i], l, paste0(tmp, make_header_latex(doc)))
```

```{r}
b = knitr::all_patterns$tex$header.begin
doc = c("\\documentclass[opt]{article}", "some tex content")
l = stringr::str_locate(doc[i], b)
```

```{r}
res = bench::mark(
  stringr = stringr::str_sub(doc[i], l[, 1], l[, 2]),
  new = substr(doc[i], l[, 1], l[, 2]),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

```{r}
tmp = stringr::str_sub(doc[i], l[, 1], l[, 2])
header = knitr:::make_header_latex(doc)

old_fun = function(doc, l, tmp, header) {
  stringr::str_sub(doc[i], l[,1], l[,2]) = paste0(tmp, header)
  doc
}

new_fun = function(doc, l, tmp, header) {
  doc[i] = str_replace(doc[i], l, paste0(tmp, header))
  doc
}

res = bench::mark(
  stringr = old_fun(doc, l, tmp, header),
  new = new_fun(doc, l, tmp, header),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```

## `stringr::str_detect()`

| Commit                                                                                    | PR                                                                                                                                                            | File                                                                                                          |
|-------------------------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------------------|
| [1a0f2cc](https://github.com/yihui/knitr/commit/1a0f2cc269c6a7f86f10b77d12229a80c14d6ee6) | [\#](https://github.com/yihui/knitr/pull/2187)[pattern.R#L155](https://github.com/yihui/knitr/blob/1a0f2cc269c6a7f86f10b77d12229a80c14d6ee6/R/pattern.R#L155) | [pandoc.R#L125](https://github.com/yihui/knitr/blob/67b973dc87c9d479cd5fbd76806e3a91f5e7d18c/R/pandoc.R#L125) |

Before:

``` r
if (length(pat) && any(stringr::str_detect(text, pat))) return(p)
```

After:

``` r
if (length(pat) && any(grepl(pat, text, perl = TRUE))) return(p)
```

```{r}
pat = knitr:::all_patterns$md$chunk.begin
text = xfun::read_utf8(knitr_example("knitr-spin.Rmd"))
```

```{r}
res = bench::mark(
  stringr = stringr::str_detect(text, pat),
  new = grepl(pat, text, perl = TRUE),
  min_time = Inf
)
summary(res, relative = TRUE)
ggplot2::autoplot(res)
```
